---
title: "Restaurant Revenue Prediction"
author: "Trevor Okinda"
date: "2024"
output:
  github_document: 
    toc: yes
    toc_depth: 4
    fig_width: 6
    fig_height: 4
    df_print: default
editor_options:
  chunk_output_type: console
---

# Student Details

|                                              |     |
|----------------------------------------------|-----|
| **Student ID Number**                        | 134780 |
| **Student Name**                             | Trevor Okinda |
| **BBIT 4.2 Group**                           | C |
| **Project Name**                             | Restaurant Revenue Prediction |

# Setup Chunk

**Note:** the following KnitR options have been set as the global defaults: <BR> `knitr::opts_chunk$set(echo = TRUE, warning = FALSE, eval = TRUE, collapse = FALSE, tidy = TRUE)`.

More KnitR options are documented here <https://bookdown.org/yihui/rmarkdown-cookbook/chunk-options.html> and here <https://yihui.org/knitr/options/>.

```{r setup, include=FALSE}
library(formatR)
knitr::opts_chunk$set(
  warning = FALSE,
  collapse = FALSE
)
```

### Source: 

The dataset that was used can be downloaded here: *\<https://www.kaggle.com/datasets/mrsimple07/restaurants-revenue-prediction\>*

### Reference:

*\<MrSimple07. (2024). Restaurants Revenue Prediction [Dataset]. Kaggle. https://www.kaggle.com/datasets/mrsimple07/restaurants-revenue-prediction\>\
Refer to the APA 7th edition manual for rules on how to cite datasets: <https://apastyle.apa.org/style-grammar-guidelines/references/examples/data-set-references>*

# Exploratory Data Analysis
## Load dataset
```{r Load dataset}
# Load necessary libraries
library(tidyverse)

# Load dataset
restaurant_data <- read.csv("Restaurant_revenue.csv", colClasses = c(
  Number_of_Customers = "integer",
  Menu_Price = "numeric",
  Marketing_Spend = "numeric",
  Cuisine_Type = "factor",
  Average_Customer_Spending = "numeric",
  Promotions = "factor",
  Reviews = "numeric",
  Monthly_Revenue = "numeric"
))

# Display the structure of the dataset
str(restaurant_data)
```

## Measures of Frequency
```{r MOF}
# Load necessary libraries
library(tidyverse)
library(psych)  # for descriptive statistics
library(GGally)  # for correlation plots

# Measures of Frequency
# Frequency of factor variables
frequency_promotions <- table(restaurant_data$Promotions)
frequency_cuisine <- table(restaurant_data$Cuisine_Type)
print("Frequency of Promotions:")
print(frequency_promotions)
print("Frequency of Cuisine Type:")
print(frequency_cuisine)
```

## Measures of Central Tendency
```{r MOCT}
# Measures of Central Tendency
# Mean, Median, Mode
mean_revenue <- mean(restaurant_data$Monthly_Revenue)
median_revenue <- median(restaurant_data$Monthly_Revenue)
mode_revenue <- as.numeric(names(sort(table(restaurant_data$Monthly_Revenue), decreasing = TRUE)[1]))

print(paste("Mean Monthly Revenue:", mean_revenue))
print(paste("Median Monthly Revenue:", median_revenue))
print(paste("Mode Monthly Revenue:", mode_revenue))
```

## Measures of Distribution
```{r MOD}
# Measures of Distribution
# Range, Variance, Standard Deviation, Skewness, Kurtosis
range_revenue <- range(restaurant_data$Monthly_Revenue)
variance_revenue <- var(restaurant_data$Monthly_Revenue)
sd_revenue <- sd(restaurant_data$Monthly_Revenue)
skewness_revenue <- skew(restaurant_data$Monthly_Revenue)
kurtosis_revenue <- kurtosi(restaurant_data$Monthly_Revenue)

print(paste("Range of Monthly Revenue:", paste(range_revenue, collapse = " - ")))
print(paste("Variance of Monthly Revenue:", variance_revenue))
print(paste("Standard Deviation of Monthly Revenue:", sd_revenue))
print(paste("Skewness of Monthly Revenue:", skewness_revenue))
print(paste("Kurtosis of Monthly Revenue:", kurtosis_revenue))
```

## Measures of Relationship
```{r MOR}
# Measures of Relationship
# Correlation matrix
correlation_matrix <- cor(restaurant_data %>% select_if(is.numeric))

print("Correlation Matrix:")
print(correlation_matrix)

# Visualize the correlation matrix
ggcorr(restaurant_data %>% select_if(is.numeric), label = TRUE)
```

## ANOVA
```{r ANOVA}
# Load necessary libraries
library(tidyverse)
library(car)  # for ANOVA assumptions tests
library(ggplot2)

# ANOVA to compare Monthly Revenue across different Cuisine Types
anova_cuisine <- aov(Monthly_Revenue ~ Cuisine_Type, data = restaurant_data)
summary(anova_cuisine)

# ANOVA to compare Monthly Revenue across different Promotions
anova_promotions <- aov(Monthly_Revenue ~ Promotions, data = restaurant_data)
summary(anova_promotions)

# Check assumptions of ANOVA

# 1. Homogeneity of variances
leveneTest(Monthly_Revenue ~ Cuisine_Type, data = restaurant_data)
leveneTest(Monthly_Revenue ~ Promotions, data = restaurant_data)

# 2. Normality of residuals
# Residual plots for Cuisine_Type ANOVA
par(mfrow = c(2, 2))
plot(anova_cuisine)

# Residual plots for Promotions ANOVA
par(mfrow = c(2, 2))
plot(anova_promotions)
```

## Plots
```{r Plots}
# Visualize the results

# Boxplot for Cuisine_Type
ggplot(restaurant_data, aes(x = Cuisine_Type, y = Monthly_Revenue)) +
  geom_boxplot() +
  labs(title = "Monthly Revenue by Cuisine Type", x = "Cuisine Type", y = "Monthly Revenue")

# Boxplot for Promotions
ggplot(restaurant_data, aes(x = Promotions, y = Monthly_Revenue)) +
  geom_boxplot() +
  labs(title = "Monthly Revenue by Promotions", x = "Promotions", y = "Monthly Revenue")

# Load necessary libraries
library(tidyverse)


# Univariate Plots

# Histogram for Monthly Revenue
ggplot(restaurant_data, aes(x = Monthly_Revenue)) +
  geom_histogram(binwidth = 50, fill = "blue", color = "black") +
  labs(title = "Histogram of Monthly Revenue", x = "Monthly Revenue", y = "Frequency")

# Density plot for Monthly Revenue
ggplot(restaurant_data, aes(x = Monthly_Revenue)) +
  geom_density(fill = "blue", alpha = 0.5) +
  labs(title = "Density Plot of Monthly Revenue", x = "Monthly Revenue", y = "Density")

# Boxplot for Monthly Revenue
ggplot(restaurant_data, aes(y = Monthly_Revenue)) +
  geom_boxplot(fill = "blue") +
  labs(title = "Boxplot of Monthly Revenue", y = "Monthly Revenue")

# Bar plot for Cuisine Type
ggplot(restaurant_data, aes(x = Cuisine_Type)) +
  geom_bar(fill = "blue") +
  labs(title = "Bar Plot of Cuisine Type", x = "Cuisine Type", y = "Count")

# Bar plot for Promotions
ggplot(restaurant_data, aes(x = Promotions)) +
  geom_bar(fill = "blue") +
  labs(title = "Bar Plot of Promotions", x = "Promotions", y = "Count")

# Multivariate Plots

# Scatter plot for Average Customer Spending vs Monthly Revenue
ggplot(restaurant_data, aes(x = Average_Customer_Spending, y = Monthly_Revenue)) +
  geom_point(color = "blue") +
  labs(title = "Scatter Plot of Average Customer Spending vs Monthly Revenue",
       x = "Average Customer Spending", y = "Monthly Revenue")

# Scatter plot for Marketing Spend vs Monthly Revenue, colored by Cuisine Type
ggplot(restaurant_data, aes(x = Marketing_Spend, y = Monthly_Revenue, color = Cuisine_Type)) +
  geom_point() +
  labs(title = "Scatter Plot of Marketing Spend vs Monthly Revenue by Cuisine Type",
       x = "Marketing Spend", y = "Monthly Revenue")

# Boxplot of Monthly Revenue by Cuisine Type
ggplot(restaurant_data, aes(x = Cuisine_Type, y = Monthly_Revenue)) +
  geom_boxplot(fill = "blue") +
  labs(title = "Boxplot of Monthly Revenue by Cuisine Type", x = "Cuisine Type", y = "Monthly Revenue")

# Boxplot of Monthly Revenue by Promotions
ggplot(restaurant_data, aes(x = Promotions, y = Monthly_Revenue)) +
  geom_boxplot(fill = "blue") +
  labs(title = "Boxplot of Monthly Revenue by Promotions", x = "Promotions", y = "Monthly Revenue")

# Pair plot for numeric variables
ggpairs(restaurant_data %>% select_if(is.numeric), title = "Pair Plot of Numeric Variables")

```

# Preprocessing and Data Transformation
## Missing Values
```{r Missing Values}
# Confirm presence of missing values

# Summary of missing values for each column
missing_values_summary <- colSums(is.na(restaurant_data))
print("Summary of missing values for each column:")
print(missing_values_summary)

# Total number of missing values in the dataset
total_missing_values <- sum(is.na(restaurant_data))
print(paste("Total number of missing values in the dataset:", total_missing_values))
```

## Round-Off Money Columns
```{r Round-Off Data}
# Round off specified columns to 2 decimal places
restaurant_data <- restaurant_data %>%
  mutate(
    Monthly_Revenue = round(Monthly_Revenue, 2),
    Average_Customer_Spending = round(Average_Customer_Spending, 2),
    Marketing_Spend = round(Marketing_Spend, 2),
    Menu_Price = round(Menu_Price, 2)
  )


# Display the first few rows of the dataset to verify the rounding
head(restaurant_data)

# Display the structure of the dataset to verify the data type
str(restaurant_data)

# View the dataset in a separate viewer window
View(restaurant_data)
```

# Model Training
## Data Splitting
```{r Data Splitting}
# Load necessary libraries
library(caTools)  # For data splitting

# Split the data into training and testing sets
set.seed(123)  # Set seed for reproducibility
split <- sample.split(restaurant_data$Monthly_Revenue, SplitRatio = 0.7)
train_data <- subset(restaurant_data, split == TRUE)
test_data <- subset(restaurant_data, split == FALSE)

# Summary of the split
cat("Training set size:", nrow(train_data), "\n")
cat("Testing set size:", nrow(test_data), "\n")
```

## Bootstrapping
```{r Bootstrapping}
# Load necessary libraries
library(boot)  # For bootstrapping

# Define the function to compute the statistic of interest (e.g., mean)
bootstrap_mean <- function(data, indices) {
  sample_data <- data[indices, ]
  return(mean(sample_data$Monthly_Revenue))
}

# Perform bootstrap resampling
set.seed(123)  # Set seed for reproducibility
boot_results <- boot(restaurant_data, statistic = bootstrap_mean, R = 1000)

# Summary of bootstrap results
summary(boot_results)
```

## Cross-validation
```{r Cross-validation}
# Load necessary libraries
library(caret)  # For cross-validation

# Define the control parameters for cross-validation
ctrl <- trainControl(method = "cv", number = 5)  # 5-fold cross-validation

# Train a linear regression model using cross-validation
set.seed(123)  # Set seed for reproducibility
lm_model <- train(Monthly_Revenue ~ ., data = train_data, method = "lm", trControl = ctrl)

# Summary of the cross-validated linear regression model
print(lm_model)
```

## Model Training
```{r Model Training}
# Load necessary libraries
library(caret)       # For model training and evaluation
library(rpart)       # For decision trees
library(randomForest)# For random forest

# Define the control parameters for cross-validation
ctrl <- trainControl(method = "cv", number = 5)  # 5-fold cross-validation

# Train a linear regression model
set.seed(123)  # Set seed for reproducibility
lm_model <- train(Monthly_Revenue ~ ., data = train_data, method = "lm", trControl = ctrl)

# Train a decision tree model
set.seed(123)  # Set seed for reproducibility
dt_model <- train(Monthly_Revenue ~ ., data = train_data, method = "rpart", trControl = ctrl)

# Train a random forest model
set.seed(123)  # Set seed for reproducibility
rf_model <- train(Monthly_Revenue ~ ., data = train_data, method = "rf", trControl = ctrl)
```

## Performance Comparison
```{r Performance Comparison}
# Summary of the models
print(lm_model)
print(dt_model)
print(rf_model)

# Compare model performance using resamples
models <- list("Linear Regression" = lm_model,
               "Decision Tree" = dt_model,
               "Random Forest" = rf_model)
resamples <- resamples(models)

# Summarize the resampling results
summary(resamples)
```

# Saving Model
```{r Saving Model}
# Saving the Linear Regression (lm) model
saveRDS(lm_model, "./models/saved_lm_model.rds")

# Load the saved model
loaded_lm_model <- readRDS("./models/saved_lm_model.rds")

# Prepare new data for prediction
new_data <- data.frame(
  Number_of_Customers = 61,
  Menu_Price = 43.11763549,
  Marketing_Spend = 12.66379252,
  Cuisine_Type = "Japanese",
  Average_Customer_Spending = 36.23613252,
  Promotions = "0",
  Reviews = 45
)

# Use the loaded model to make predictions
predictions_loaded_model <- predict(loaded_lm_model, newdata = new_data)

# Print predictions
print(predictions_loaded_model)

```


